Dealing with Consumer Uncertainty: Online Consumer Reviews and What makes them helpful?

### 2022-11-03 Paper.Rmd
extend intro, lit_rev, theory

However, the increased rate of electronic commerce activity also created an abundance of product reviews on the internet. The vast amount of product reviews might increase the information overload on the customer [@sun2019; @fan2021]. Unfortunately, this information overload can make the search process even costlier because it is really time consuming to evaluate thousand of customer reviews. Additionally, not all reviews have the same informativeness level [@pan2011]. Therefore, answering the question of "what makes a helpful review?" is very important because reading more and more reviews can only increase the search cost. To make the understanding of a "helpful review" more clear it can be defined as "*a peer-generated product evaluation that facilitates the consumer's purchase decision process*" [@mudambi2010].

@@

The research on perceived helpfulness is still growing. In Social and Behavioral Sciences, it is common to have conflicting papers and results in the literature [@hunter1991]. In this case, the study on perceived helpfulness is not an exception. Because of this reason, more empirical studies on perceived helpfulness are required. This paper aims to add one more perspective to the growing literature of customer reviews and review helpfulness.

This study uses secondary data. The reviews were collected from two websites. Namely, IMDB and BestBuy. The data was scrapped with Selenium Webdriver for Python [@python3]. Over 400,000 reviews and their meta data were gathered initially. After the data transformation process the final sample size was over 23,000. Then this data was subjected to empirical tests. It was found out that consumers usually find extreme reviews more helpful than neutral ones. Furthermore, consumers prefer longer reviews when reviews are rated negatively. On the contrast, positive reviews with fewer words were deemed more helpful by readers on the average. These effects were further moderated by product types, namely, Search and Experience goods.

The plan of this study is outlined as the following. In the next section, the literature on the subject of Review Helpfulness is going to be presented. In the third section, the related theories which would make the understanding of the consumer behavior and review helpfulness will be laid out. Next, the data and the research methodology are going to be introduced. The results will be shared in the fifth section. In the last section the results are going to be discussed and the study is going to be concluded.

@@

@@

Consumers can choose from two different strategies when they want to acquire more information about products. Consumers "search" in order to get information about a product assuming that they perfectly know where they can obtain the knowledge. @nelson1970 defines the search process as *"The most obvious procedure available to the consumer in obtaining information about price or quality is search"*. The search process is complete when consumers inspect the option and that inspection happens before the purchase. Contrary to the search process, consumer may choose to "experience" the product. The consumer follows this strategy when the search for information is either inappropriate, inefficient or costly compared to just experiencing the good. This is true for cheap goods especially. For instance, in order to find out the best chocolate in the market it's best for the consumer to buy every other brand available and taste them. Thus, consumers will always prefer sampling or trial use given the fact that the cost of the sampling or trial is lower than the cost of the search. This is why search is costlier for Experience goods. Therefore, the kind of strategy that a consumer chooses depends on the product's Search or Experience attributes and how dominant these attributes are.

@@

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. Someone else's hair in one's meal has the potential of ruining the whole meal. On the contrary a plate of hair will not turn to a delicious meal just because there is a slice of a delicious pizza on top of the hair. @rozin2001 explains the Negativity Bias Theorem by describing it with four main aspects. Namely, Negative Potency, Steeper Negative Gradients, Negative Dominance and Negative Differentiation [@rozin2001; @sen2007; @eslami2018]. Loss Aversion has a strong connection to Negativity Bias Theorem since negative cues might be used as an early warning against potential losses [@herr1991].

@@

**Negativity Dominance** asserts that a balanced mixture of positive and negative evaluations creates a perception of negativity. This means that negative entities have higher weight than positive entities. This is apparent in religions. For the most of the religions being a good and pious person takes time, demands utmost care. However, tainting this image is very easy with certain actions such as murder or pre-martial intercourse.

>>>>

However, the increased rate of electronic commerce activity also created an abundance of product reviews on the internet. The vast amount of product reviews might increase the information overload on the customer [@sun2019; @fan2021]. Unfortunately, this information overload can make the search process even costlier because it is really time consuming to evaluate thousand of consumer reviews. Additionally, not all reviews have the same informativeness level [@pan2011]. Therefore, answering the question of "what makes a helpful review?" is very important because reading more and more reviews without getting useful information can only increase the search cost. To make the understanding of a "helpful review" more clear it can be defined as "*a peer-generated product evaluation that facilitates the consumer's purchase decision process*" [@mudambi2010].

Having a good understanding of how consumers handle information and how their behaviors change with respect to information they are exposed to might even unearth newer tools to increase the social welfare because it would lead to decreased level of information asymmetry. The decreased information asymmetry might produce two benefits. First, consumers can learn to evaluate the quality of the products before purchase much more efficiently. This leaves less room for any dis-utility that can arise from purchasing a sub-par quality product. The second, the increased knowledge of the consumers towards certain products may even facilitate competition among firms. This competition may lead increased quality of the products as well as decreased prices in a given market. This is why understanding the determinants of a helpful review is important because it might reveal new insights on consumer behavior and these new insights could be used in favor of both consumers are producers.

@@

The research on perceived helpfulness is still growing. In Social and Behavioral Sciences, it is common to have conflicting papers and results in the literature [@hunter1991]. In this case, the study on perceived helpfulness is not an exception. Because of this reason, more empirical studies on perceived helpfulness are required. This paper aims to add one more perspective to the growing literature of customer reviews and review helpfulness by employing a unique data set as well as questioning the subject with the intuition of economics. Furthermore, this paper tries to show that unstructured data (e.g., text, image, etc.) is viable for economic research. With unstructured data, economics may be able to explore new venues.

This study uses secondary data. The reviews were collected from two websites. Namely, IMDB and BestBuy. The data was scrapped with Selenium Webdriver for Python [@python3]. Over 400,000 reviews and their meta data were gathered initially. After the data transformation process the final sample size was over 23,000. Then this data was subjected to empirical tests. It was found out that consumers usually find extreme reviews more helpful than neutral ones. Furthermore, consumers prefer longer reviews when reviews are rated negatively. On the contrast, positive reviews with fewer words were deemed more helpful by readers on the average. In addition, the sentiment of a review has also a significant effect on the helpfulness of a review. These effects were further moderated by product types, namely, Search and Experience goods.

The plan of this study is outlined as the following. In the next section, the literature on the subject of Review Helpfulness is going to be presented. In the second section, the related theories which would help with the understanding of the consumer behavior and review helpfulness will be laid out. Next, the data and the research methodology are going to be introduced. The results will be shared in the fourth section. In the last section the results are going to be discussed and then the study is going to be concluded.

@@

Since there are many conflicting results in the literature of review helpfulness additional empirical contributions are important. As well as adding new empirical findings to the literature, this study tries to extend the potential predictors with the interaction term between review length and review rating to test for the existence of the Negativity Bias. Furthermore, some under-utilized predictors from the literature such as the proportion of negative words and the proportion of positive words are also used to test Negativity Bias and Loss Aversion.

@@

Consumers can choose from two different strategies when they want to acquire more information about products. Consumers "search" in order to get information about a product assuming that they perfectly know where they can obtain the knowledge. @nelson1970 defines the search process as *"The most obvious procedure available to the consumer in obtaining information about price or quality is search"*. The search process is complete when consumers inspect the option and that inspection happens before the purchase. Contrary to the search process, consumer may choose to "experience" the product. The consumer follows this strategy when the search for information is either inappropriate, inefficient or costly compared to just experiencing the good. This is true for cheap goods especially. For instance, in order to find out the best chocolate in the market it is best for the consumer to buy every other brand available and taste them. Thus, consumers will always prefer sampling or trial use given the fact that the cost of the sampling or trial is lower than the cost of the search. This is why search is costlier for Experience goods. Therefore, the kind of strategy that a consumer chooses depends on the product's Search or Experience attributes and how dominant these attributes are.

@@

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. In @rozin2001 's words Negativity Bias can be best exemplified such as following:

> *"Brief contact with a cockroach will usually render a delicious meal inedible. The inverse phenomenon rendering a pile of cockroaches on a platter edible by contact with one's favorite food-is unheard of. More modestly, consider a dish of a food that you are inclined to dislike: lima beans, fish, or whatever. What could you touch to that food to make it desirable to eat- that is, what is the anti-cockroach? Nothing! And the cockroach is far from unique: there is a wide variety of animals (e.g., worms, caterpillars, slugs, spiders) that share the cockroach potency, along with a variety of microbially or toxin-contaminated objects. One of the best generic descriptions of this relative power of negative contamination is embedded in an age old Russian adage: 'A spoonful of tar can spoil a barrel of honey, but a spoonful of honey does nothing for a barrel of tar.'"*

@rozin2001 explains the Negativity Bias Theorem by describing it with four main aspects. Namely, Negative Potency, Steeper Negative Gradients, Negative Dominance and Negative Differentiation [@rozin2001; @sen2007; @eslami2018]. Loss Aversion has a strong connection to Negativity Bias Theorem since negative cues might be used as an early warning against potential losses [@herr1991].

@@

**Negativity Dominance** asserts that a balanced mixture of positive and negative evaluations creates a perception of negativity. This means that negative entities have higher weight than positive entities. @rozin2001 also puts forwards as an example:

> *''The image that comes to mind from the physical domain is that of a single cancerous growth or germ that radiates itself through and ultimately consumes a perfectly healthy body. The image that comes to mind from the moral domain is that of a single vice corrupting and perverting and bringing the moral downfall of an otherwise perfectly good person.''*
